---
title: "Practical 1"
author: "Jumping Rivers"
header-includes:
  - \usepackage{../inst/jrStyle}

output:
  bookdown::tufte_book2:
    latex_engine: xelatex
    toc: false
    highlight: tango
classoption: a4paper,justified,openany
vignette: >
  %\VignetteIndexEntry{Practical 1}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r include = FALSE}
jrRSConnect:::make_preamble()
```

# Getting Started

## Accessing the training machine

\sloppy
To help facilitate this practical we have provisioned a Virtual Machine (VM)
for us all to work from. The URL of this machine will be of the form
`id.jumpingrivers.training`. Your trainer will circulate this URL during the
session.


\indent Once you have the URL, visit the VM's welcome page (the circulated URL
suffixed with `/welcome`, and so a url of the form
`id.jumpingrivers.training/welcome`). Enter your email address and the supplied
master password, as illustrated in \cref{fig:reveal-login}. This will reveal a
username and password for you to log-in to the training enviornment.


\indent If you lose your username and password at any point during the
training: don't worry, re-entering your email address and the master password
on the welcome page will retrieve your login for the training machine.

\begin{marginfigure}[-8.75cm]
  \centering
  \includegraphics[width=\linewidth]{../inst/reveal-login.png}
  \captionof{figure}{Revealing login details on the welcome page.}
  \label{fig:reveal-login}
\end{marginfigure}

\indent Head over to the training enviornment---either with the circulated URL,
or by clicking the "Training Environment" link visible in the welcome page
(\cref{fig:reveal-login}). Here you will be prompted for log in details
(\cref{fig:training-login}). Use the details that were revealed to you on the
welcome page.

\begin{marginfigure}[-2cm]
  \centering
  \includegraphics[width=\linewidth]{../inst/training-login.png}
  \captionof{figure}{Entering our revealed login details on the training
  machine.}
  \label{fig:training-login}
\end{marginfigure}

\indent After authenticating with the training machine you will enter a
homepage, at which you will be able to start a new RStudio session, as
demonstrated in \cref{fig:new-session,fig:new-rstudio}. Having done so you will
be rewarded with the familiar sight of an RStudio IDE.

\begin{marginfigure}
  \centering
  \includegraphics[width=\linewidth]{../inst/new-session.png}
  \captionof{figure}{Starting a new session within the training VM.}
  \label{fig:new-session}
\end{marginfigure}

\begin{marginfigure}
  \centering
  \includegraphics[width=\linewidth]{../inst/new-rstudio.png}
  \captionof{figure}{Instantiating an RStudio session within the VM.}
  \label{fig:new-rstudio}
\end{marginfigure}

## Accessing our test RStudio Connect

Along with the training machine we have deployed a sandpit RStudio Connect
service for us to have a play around in. This Connect service is accessible as
the training URL, appended with `/connect`: so a URL of the form
`id.jumpingrivers.training/connect`.

Use your username and login, as revealed in \cref{fig:reveal-login}, to
authenticate against this Connect server.

# Publishing Manually: Hello, User!

In this example we are going to walk through the steps of publishing an `Rmd`
file from RStudio.

\marginpar{You can create an empty R Markdown file with the \texttt{Create
Empty Document} button in the bottom left of the \texttt{New R Markdown}
window.}

1. Create a new `Rmd` file in RStudio with `File -> New File -> Rmarkdown`.

2. Into this new file paste the following code, being mindful to preserve any
   indentation (particularly in the yaml front matter):

\begin{shaded}
\begin{verbatim}
`r paste(readLines('../inst/hello-user.Rmd'), collapse = '\n')`
\end{verbatim}
\end{shaded}

3. Verify that you can render this `Rmd` with the knit button---if you haven't
   done so already, at this point you will be prompted to save your file.

\begin{marginfigure}
  \centering
  \includegraphics[width=0.5\linewidth]{../inst/knit.png}
  \captionof{figure}{The knit button in the RStudio IDE.}
  \label{fig:knit}
\end{marginfigure}

4. Click the blue publish icon in the Connect IDE. Follow the steps to publish
   the content to the Connect server associated with this training machine.
   The Connect server's public URL is the same as the URL of the training
   machine (of the form `id.jumpingrivers.training`). You will need your
   username uXXX and its corresponding password to perform this step.

5. Log in to the Connect service with your username and password and verify that
   you can access your report.

# Detecting the current user

In this example we are going to publish a shiny application to our Connect
server which accesses the username of the content viewer, and displays output
to this user conditional on their username.

1. Create a new shiny app in RStudio with `File -> New File -> Shiny Web App`.
   Choose an application name, and choose "Single File (app.R)" for the
   application type.

2. Delete the pre-populated content of `app.R` and paste in the following
   code:

```{r eval=FALSE, tidy=FALSE, code.cap="Code to create a shiny app that greets the current user.", label="shiny"}
library("shiny")

# Define UI for app
ui <- fluidPage(
  # App title ----
  titlePanel("Shiny ID"),
  textOutput("welcome")
)

# Define server logic required to draw a histogram
server <- function(input, output, session) {
  client_email_address = session$user

  output$welcome <- renderText({
    paste0("Hi, ", client_email_address, ".")
  })

}

# Run the application
shinyApp(ui = ui, server = server)
```

3. Click the blue publish button in the RStudio IDE and follow the steps to
   deploy this content to Connect. Verify that you can access this content and
   that it correctly displays your username.

4. Alter your app code as copied from \cref{chk:shiny} to conitionally display
   different output for your user. Here's one I made earlier (though try not to
   just copy and paste this example!).

```{r eval=FALSE, code.cap="Altering the code in \\cref{chk:shiny} to conditionally show output based on the current user."}
library("shiny")

# Define UI for app
ui <- fluidPage(
    # App title ----
    titlePanel("Shiny ID"),
    textOutput("welcome"),
    imageOutput("image")
)

# Define server logic required to draw a histogram
server <- function(input, output, session) {
    client_email_address = session$user

    output$welcome <- renderText({
        paste0("Hi, ", client_email_address, ".")
    })

    output$image = renderImage({
        outfile <- tempfile(fileext = ".png")
        # Generate a png
        png(outfile,
            width = 400,
            height = 400)

        if (client_email_address == "u047") {
            hist(rnorm(100))
        } else {
            hist(runif(100))
        }
        dev.off()

        # Return a list
        list(src = outfile,
             alt = "Alt text")
        }, deleteFile = TRUE)
}

# Run the application
shinyApp(ui = ui, server = server)
```
